Determining the alternatives for scalar implicature

### 2016-01-27 Paper/Paper.Rmd
new opening sentence and convered scalar items to italics from quotes

# Experiments E8, E10, E12: Literal listener
# ------------------------------------------
speakerE12 <- read.csv("../models/model_data/L0_e12.csv", stringsAsFactors = FALSE)
speakerE10 <- read.csv("../models/model_data/L0_e10.csv", stringsAsFactors = FALSE)
# speakerE8 <- read.csv("../models/model_data/L0_e8.csv", stringsAsFactors = FALSE)
speakerE8 <- read.csv("../models/model_data/L0_e8Compliance.csv", stringsAsFactors = FALSE)
# Experiments E6, E11: Pragmatic listener
# ---------------------------------------
listenerE6 <- read.csv("../models/model_data/L1_e6.csv", stringsAsFactors = FALSE)
listenerE11 <- read.csv("../models/model_data/L1_e11.csv", stringsAsFactors = FALSE)
```

# Humpty Dumpty Quote...
# Introduction

Successful communication requires listeners to make pragmatic inferences that go beyond the literal semantic content of speakers' utterances. For example, listeners  commonly enrich the meaning of the scalar item "some" to "some but not all" in sentences like "Bob ate some of the cookies" [@grice1975;@horn1984;@levinson2000]. These inferences, called _scalar implicatures_, have been an important test case for understanding pragmatic inferences more generally. A Gricean account of this phenomena assumes listeners reason about the meaning the speaker intended by incorporating knowledge about a) alternative scalar items a speaker could have used (such as "all") and b) the relative informativity of using such alternatives [@grice1975]. According to this account, a listener will infer that the speaker must have intended that Bob did not eat "all" the cookies because it would have been *underinformative* for the speaker to use "some" when "all" could have been used.

But what are the alternatives that should be considered in this computation more generally? Under classic accounts of implicature, listeners consider only those words whose meaning would entail the word that is actually sent [@horn1972], and these alternatives enter into conventionalized or semi-conventionalized scales [@levinson2000]. For example, because "all" entails "some," and hence is a "stronger" meaning, "all" should be considered as an alternative to "some" in implicatures. Similar scales exist for non-quantifier scales, e.g. "love" entails "like" (and hence "I liked the movie" implicates that I didn't love it).

Recent empirical evidence has called into question whether entailment scales are all that is necessary for understanding scalar implicature. For example, @degen2015  demonstrated that the scalar item "some" was judged less appropriate when exact numbers were seen as viable alternatives. And in a different paradigm, @vantiel2014 found converging evidence that "some" was judged to be atypical for small quantities. These data provide indirect evidence about a broader set of alternatives: since "some" is logically true of sets with one or two members, these authors argued that the presence of salient alternatives (the words "one" and "two") reduced the felicity of "some" via a pragmatic inference.

By formalizing pragmatic reasoning, computational models can help provide more direct evidence about the role that alternatives play. The "rational speech act" model (RSA) is one recent framework for understanding inferences about meaning in context [@frank2012;@goodman2013]. RSA models frame language understanding as a special case of social cognition, in which listeners and speakers reason recursively about one another's goals. In the case of scalar implicature, a listener makes a probabilistic inference about what the speaker's most likely communicative goal was, given that she picked the quantifier "some" rather than the stronger quantifier "all." In turn, the speaker reasons about what message would best convey her intended meaning to the listener, given that he is reasoning in this way. This recursion is grounded in a "literal" listener who reasons only according to the basic truth-functional semantics of the language.

@franke2014 used an RSA-style model to assess what alternatives a speaker would need to consider in order to produce the typicality/felicity ratings reported by @degen2015 and @vantiel2014. In order to do this, @franke2014's model assigned weights to a set of alternative numerical expressions. Surprisingly, along with weighting "one" highly (a conclusion that was suppored by the empirical work), the best-fitting model assigned substantial weight to "none" as an alternative. This finding was especially surprising considering the emphasis of standard theories on scalar items that stand in entailment relationships with one another (e.g. "one" entails "some" even if it is not classically considered to be part of the scale).

In our current work, we pick up where these previous studies left off,  considering the set of alternatives for implicature using the RSA model. To gain empirical traction on this issue, however, we broaden the set of scales we consider. Our inspiration for this move comes from work by @vantiel2014b, who examined a phenomenon that they dubbed "scalar diversity," namely the substantial difference in the strength of scalar implicature across a variety of scalar pairs (e.g. "liked/loved," or "palatable/delicious."). Making use of some of this diversity allows us to investigate the ways that different alternative sets give rise to implicatures of different strengths across scales.

@@

Figure 2 plots literal listener $p_{L0}(m|u)$ distributions for the three conditions. Several trends are visible. First, in each scale, the alternatives spanned the star scale, such that there were alternatives that were highly compatible with both the lowest and highest numbers of stars. Second, there was clear variability between scalar families. For example, compatibility judgments for the top items in the "memorable / unforgettable" scale were more similar than those for "good / excellent" or "liked / loved." Finally, there was substantial consistency in ratings for items that were repeated across experiments, suggesting that this paradigm elicited stable judgments from participants.

@@

Figure \ref{fig:exp2_altsPlot} an example alternative set for the scalar items "liked" and "loved" (combined). Alternative distributions for the other scalar pairs (e.g.. "good/excellent", "memorable/unforgettable") were similarly long-tailed.

@@

In Experiment 3, we asked participants to make pragmatic listener judgements.

CONDITIONS

@@

We recruited 150 participants from AMT, 50 for each condition. Data for 7 participants in the two-alternative condition were excluded after participants either failed to pass two training trials or reported a non-English native language, leaving a total sample of 43 participants. In the four-alternative condition, data from five participants was thrown out after participants either failed to pass two training trials or were not native English speakers, leaving a total sample of 45 participants.

>>>>

# Non-summarized data
# -------------------
allLiteralListener <- read.csv("../models/model_data/RawLiteralListenerCombined.csv", 
                               stringsAsFactors = FALSE)
allPragmaticListener <- read.csv("../models/model_data/RawPragmaticListenerCombined.csv", 
                               stringsAsFactors = FALSE)

# Model run data
# --------------
# Speakers ::
speakerE12 <- read.csv("../models/model_data/L0_e12.csv", stringsAsFactors = FALSE)
speakerE10 <- read.csv("../models/model_data/L0_e10.csv", stringsAsFactors = FALSE)
# speakerE8 <- read.csv("../models/model_data/L0_e8.csv", stringsAsFactors = FALSE)
speakerE8 <- read.csv("../models/model_data/L0_e8Compliance.csv", stringsAsFactors = FALSE)
# Listeners ::
listenerE6 <- read.csv("../models/model_data/L1_e6.csv", stringsAsFactors = FALSE)
listenerE11 <- read.csv("../models/model_data/L1_e11.csv", stringsAsFactors = FALSE)

# \begin{center}
# \textit{``When I choose a word," Humpty Dumpty said, in a rather scornful tone, ``it means just what I choose it to mean - neither more nor less"- Lewis Caroll, Through the Looking-Glass}
# \end{center}
```

# Introduction

How much of what we mean comes from the words that go unsaid? As listeners, the ability to enrich meaning in context is an indispensible component of succesful communication.  For example, listeners  commonly enrich the meaning of the scalar item *some* to *some but not all* in sentences like "Alice ate some of the cookies" [@grice1975;@horn1984;@levinson2000]. These inferences, called _scalar implicatures_, have been an important test case for understanding pragmatic inferences more generally. A Gricean account of this phenomena assumes listeners reason about the meaning the speaker intended by incorporating knowledge about a) alternative scalar items a speaker could have used (such as *all*) and b) the relative informativity of using such alternatives [@grice1975]. According to this account, a listener will infer that the speaker must have intended that Alice did not eat *all* the cookies because it would have been *underinformative* for the speaker to use *some* when *all* could have been used.

But how much of the meaning of *some* comes from the alternatives that weren't chosen? Put differently, what are the alternatives that should be considered in this computation more generally? Under classic accounts of implicature, listeners consider only those words whose meaning would entail the word that is actually sent [@horn1972], and these alternatives enter into conventionalized or semi-conventionalized scales [@levinson2000]. For example, because *all* entails *some*, and hence is a "stronger" meaning, *all* should be considered as an alternative to *some* in implicatures. Similar scales exist for non-quantifier scales, e.g. *loved* entails *liked* (and hence "I liked the movie" implicates that I didn't love it).

Recent empirical evidence has called into question whether entailment scales are all that is necessary for understanding scalar implicature. For example, @degen2015  demonstrated that the scalar item *some* was judged less appropriate when exact numbers were seen as viable alternatives. And in a different paradigm, @vantiel2014 found converging evidence that *some* was judged to be atypical for small quantities. These data provide indirect evidence about a broader set of alternatives: since *some* is logically true of sets with one or two members, these authors argued that the presence of salient alternatives (the words *one* and *two*) reduced the felicity of *some* via a pragmatic inference.

By formalizing pragmatic reasoning, computational models can help provide more direct evidence about the role that alternatives play. The "rational speech act" model (RSA) is one recent framework for understanding inferences about meaning in context [@frank2012;@goodman2013]. RSA models frame language understanding as a special case of social cognition, in which listeners and speakers reason recursively about one another's goals. In the case of scalar implicature, a listener makes a probabilistic inference about what the speaker's most likely communicative goal was, given that she picked the quantifier *some* rather than the stronger quantifier *all*. In turn, the speaker reasons about what message would best convey her intended meaning to the listener, given that he is reasoning in this way. This recursion is grounded in a "literal" listener who reasons only according to the basic truth-functional semantics of the language.

@franke2014 used an RSA-style model to assess what alternatives a speaker would need to consider in order to produce the typicality/felicity ratings reported by @degen2015 and @vantiel2014. In order to do this, @franke2014's model assigned weights to a set of alternative numerical expressions. Surprisingly, along with weighting *one* highly (a conclusion that was suppored by the empirical work), the best-fitting model assigned substantial weight to *none* as an alternative. This finding was especially surprising considering the emphasis of standard theories on scalar items that stand in entailment relationships with one another (e.g. *one* entails *some* even if it is not classically considered to be part of the scale).

In our current work, we pick up where these previous studies left off,  considering the set of alternatives for implicature using the RSA model. To gain empirical traction on this issue, however, we broaden the set of scales we consider. Our inspiration for this move comes from work by @vantiel2014b, who examined a phenomenon that they dubbed "scalar diversity," namely the substantial difference in the strength of scalar implicature across a variety of scalar pairs (e.g. *liked/loved,* or *palatable/delicious.*). Making use of some of this diversity allows us to investigate the ways that different alternative sets give rise to implicatures of different strengths across scales.

@@

Figure 2 plots literal listener $p_{L0}(m|u)$ distributions for the three conditions. Several trends are visible. First, in each scale, the alternatives spanned the star scale, such that there were alternatives that were highly compatible with both the lowest and highest numbers of stars. Second, there was clear variability between scalar families. For example, compatibility judgments for the top items in the *memorable / unforgettable* scale were more similar than those for *good / excellent* or *liked / loved.* Finally, there was substantial consistency in ratings for items that were repeated across experiments, suggesting that this paradigm elicited stable judgments from participants.

@@

Figure \ref{fig:exp2_altsPlot} an example alternative set for the scalar items *liked* and *loved* (combined). Alternative distributions for the other scalar pairs (e.g.. *good/excellent*, *memorable/unforgettable*) were similarly long-tailed.

@@

Experiment 3 was conducted to measure pragmatic judgments $p_{L_1}(m \mid u)$ for five pairs of scalar items taken from @vantiel2014. We include two conditions in Experiment 3. In the two-alternative entailment condition, participants made judgments for items included in the entailment scales. In the four-alternative condition, participants made pragmatic judgments for the entialment items and also the top two alternatives elicited for each scale in Experiment 2. Running two pragmatic judgment conditions allowed us to rule out the potential effects of having a larger set of alternatives during the pragmatic judgment elicitation and also provided two sets of human judgments to compare with model predictions. \footnote{Note that alternatives in the four-alternative condition were chosen on the basis of Experiment 2, which was run chronologically after the two-alternative condition; both pragmatic listener experiments are grouped together for simplicity in reporting.}

@@

We recruited 100 participants from AMT, 50 for each condition. Data for 7 participants in the two-alternative condition were excluded after participants either failed to pass two training trials or reported a non-English native language, leaving a total sample of 43 participants. In the four-alternative condition, data from five participants was thrown out after participants either failed to pass two training trials or were not native English speakers, leaving a total sample of 45 participants.

### 2016-01-27 Paper/Paper.tex
new opening sentence and convered scalar items to italics from quotes

\section{Humpty Dumpty Quote\ldots{}}\label{humpty-dumpty-quote}

@@

Successful communication requires listeners to make pragmatic inferences
that go beyond the literal semantic content of speakers' utterances. For
example, listeners commonly enrich the meaning of the scalar item
``some'' to ``some but not all'' in sentences like ``Bob ate some of the
cookies'' (Grice, 1975; Horn, 1984; Levinson, 2000). These inferences,
called \emph{scalar implicatures}, have been an important test case for
understanding pragmatic inferences more generally. A Gricean account of
this phenomena assumes listeners reason about the meaning the speaker
intended by incorporating knowledge about a) alternative scalar items a
speaker could have used (such as ``all'') and b) the relative
informativity of using such alternatives (Grice, 1975). According to
this account, a listener will infer that the speaker must have intended
that Bob did not eat ``all'' the cookies because it would have been
\emph{underinformative} for the speaker to use ``some'' when ``all''
could have been used.

But what are the alternatives that should be considered in this
computation more generally? Under classic accounts of implicature,
listeners consider only those words whose meaning would entail the word
that is actually sent (Horn, 1972), and these alternatives enter into
conventionalized or semi-conventionalized scales (Levinson, 2000). For
example, because ``all'' entails ``some,'' and hence is a ``stronger''
meaning, ``all'' should be considered as an alternative to ``some'' in
implicatures. Similar scales exist for non-quantifier scales, e.g.
``love'' entails ``like'' (and hence ``I liked the movie'' implicates
that I didn't love it).

Recent empirical evidence has called into question whether entailment
scales are all that is necessary for understanding scalar implicature.
For example, Degen \& Tanenhaus (2015) demonstrated that the scalar item
``some'' was judged less appropriate when exact numbers were seen as
viable alternatives. And in a different paradigm, Tiel (2014) found
converging evidence that ``some'' was judged to be atypical for small
quantities. These data provide indirect evidence about a broader set of
alternatives: since ``some'' is logically true of sets with one or two
members, these authors argued that the presence of salient alternatives
(the words ``one'' and ``two'') reduced the felicity of ``some'' via a
pragmatic inference.

By formalizing pragmatic reasoning, computational models can help
provide more direct evidence about the role that alternatives play. The
``rational speech act'' model (RSA) is one recent framework for
understanding inferences about meaning in context (Frank \& Goodman,
2012; N. D. Goodman \& Stuhlm{ü}ller, 2013). RSA models frame language
understanding as a special case of social cognition, in which listeners
and speakers reason recursively about one another's goals. In the case
of scalar implicature, a listener makes a probabilistic inference about
what the speaker's most likely communicative goal was, given that she
picked the quantifier ``some'' rather than the stronger quantifier
``all.'' In turn, the speaker reasons about what message would best
convey her intended meaning to the listener, given that he is reasoning
in this way. This recursion is grounded in a ``literal'' listener who
reasons only according to the basic truth-functional semantics of the
language.

Franke (2014) used an RSA-style model to assess what alternatives a
speaker would need to consider in order to produce the
typicality/felicity ratings reported by Degen \& Tanenhaus (2015) and
Tiel (2014). In order to do this, Franke (2014)'s model assigned weights
to a set of alternative numerical expressions. Surprisingly, along with
weighting ``one'' highly (a conclusion that was suppored by the
empirical work), the best-fitting model assigned substantial weight to
``none'' as an alternative. This finding was especially surprising
considering the emphasis of standard theories on scalar items that stand
in entailment relationships with one another (e.g. ``one'' entails
``some'' even if it is not classically considered to be part of the
scale).

In our current work, we pick up where these previous studies left off,
considering the set of alternatives for implicature using the RSA model.
To gain empirical traction on this issue, however, we broaden the set of
scales we consider. Our inspiration for this move comes from work by Van
Tiel, Van Miltenburg, Zevakhina, \& Geurts (2014), who examined a
phenomenon that they dubbed ``scalar diversity,'' namely the substantial
difference in the strength of scalar implicature across a variety of
scalar pairs (e.g. ``liked/loved,'' or ``palatable/delicious.''). Making
use of some of this diversity allows us to investigate the ways that
different alternative sets give rise to implicatures of different
strengths across scales.

@@

Figure 2 plots literal listener \(p_{L0}(m|u)\) distributions for the
three conditions. Several trends are visible. First, in each scale, the
alternatives spanned the star scale, such that there were alternatives
that were highly compatible with both the lowest and highest numbers of
stars. Second, there was clear variability between scalar families. For
example, compatibility judgments for the top items in the ``memorable /
unforgettable'' scale were more similar than those for ``good /
excellent'' or ``liked / loved.'' Finally, there was substantial
consistency in ratings for items that were repeated across experiments,
suggesting that this paradigm elicited stable judgments from
participants.

@@

Figure \ref{fig:exp2_altsPlot} an example alternative set for the scalar
items ``liked'' and ``loved'' (combined). Alternative distributions for
the other scalar pairs (e.g.. ``good/excellent'',
``memorable/unforgettable'') were similarly long-tailed.

@@

In Experiment 3, we asked participants to make pragmatic listener
judgements.

CONDITIONS

>>>>

@@

How much of what we mean comes from the words that go unsaid? As
listeners, the ability to enrich meaning in context is an indispensible
component of succesful communication. For example, listeners commonly
enrich the meaning of the scalar item \emph{some} to \emph{some but not
all} in sentences like ``Alice ate some of the cookies'' (Grice, 1975;
Horn, 1984; Levinson, 2000). These inferences, called \emph{scalar
implicatures}, have been an important test case for understanding
pragmatic inferences more generally. A Gricean account of this phenomena
assumes listeners reason about the meaning the speaker intended by
incorporating knowledge about a) alternative scalar items a speaker
could have used (such as \emph{all}) and b) the relative informativity
of using such alternatives (Grice, 1975). According to this account, a
listener will infer that the speaker must have intended that Alice did
not eat \emph{all} the cookies because it would have been
\emph{underinformative} for the speaker to use \emph{some} when
\emph{all} could have been used.

But how much of the meaning of \emph{some} comes from the alternatives
that weren't chosen? Put differently, what are the alternatives that
should be considered in this computation more generally? Under classic
accounts of implicature, listeners consider only those words whose
meaning would entail the word that is actually sent (Horn, 1972), and
these alternatives enter into conventionalized or semi-conventionalized
scales (Levinson, 2000). For example, because \emph{all} entails
\emph{some}, and hence is a ``stronger'' meaning, \emph{all} should be
considered as an alternative to \emph{some} in implicatures. Similar
scales exist for non-quantifier scales, e.g. \emph{loved} entails
\emph{liked} (and hence ``I liked the movie'' implicates that I didn't
love it).

Recent empirical evidence has called into question whether entailment
scales are all that is necessary for understanding scalar implicature.
For example, Degen \& Tanenhaus (2015) demonstrated that the scalar item
\emph{some} was judged less appropriate when exact numbers were seen as
viable alternatives. And in a different paradigm, Tiel (2014) found
converging evidence that \emph{some} was judged to be atypical for small
quantities. These data provide indirect evidence about a broader set of
alternatives: since \emph{some} is logically true of sets with one or
two members, these authors argued that the presence of salient
alternatives (the words \emph{one} and \emph{two}) reduced the felicity
of \emph{some} via a pragmatic inference.

By formalizing pragmatic reasoning, computational models can help
provide more direct evidence about the role that alternatives play. The
``rational speech act'' model (RSA) is one recent framework for
understanding inferences about meaning in context (Frank \& Goodman,
2012; N. D. Goodman \& Stuhlm{ü}ller, 2013). RSA models frame language
understanding as a special case of social cognition, in which listeners
and speakers reason recursively about one another's goals. In the case
of scalar implicature, a listener makes a probabilistic inference about
what the speaker's most likely communicative goal was, given that she
picked the quantifier \emph{some} rather than the stronger quantifier
\emph{all}. In turn, the speaker reasons about what message would best
convey her intended meaning to the listener, given that he is reasoning
in this way. This recursion is grounded in a ``literal'' listener who
reasons only according to the basic truth-functional semantics of the
language.

Franke (2014) used an RSA-style model to assess what alternatives a
speaker would need to consider in order to produce the
typicality/felicity ratings reported by Degen \& Tanenhaus (2015) and
Tiel (2014). In order to do this, Franke (2014)'s model assigned weights
to a set of alternative numerical expressions. Surprisingly, along with
weighting \emph{one} highly (a conclusion that was suppored by the
empirical work), the best-fitting model assigned substantial weight to
\emph{none} as an alternative. This finding was especially surprising
considering the emphasis of standard theories on scalar items that stand
in entailment relationships with one another (e.g. \emph{one} entails
\emph{some} even if it is not classically considered to be part of the
scale).

In our current work, we pick up where these previous studies left off,
considering the set of alternatives for implicature using the RSA model.
To gain empirical traction on this issue, however, we broaden the set of
scales we consider. Our inspiration for this move comes from work by Van
Tiel, Van Miltenburg, Zevakhina, \& Geurts (2014), who examined a
phenomenon that they dubbed ``scalar diversity,'' namely the substantial
difference in the strength of scalar implicature across a variety of
scalar pairs (e.g. \emph{liked/loved,} or \emph{palatable/delicious.}).
Making use of some of this diversity allows us to investigate the ways
that different alternative sets give rise to implicatures of different
strengths across scales.

@@

Figure 2 plots literal listener \(p_{L0}(m|u)\) distributions for the
three conditions. Several trends are visible. First, in each scale, the
alternatives spanned the star scale, such that there were alternatives
that were highly compatible with both the lowest and highest numbers of
stars. Second, there was clear variability between scalar families. For
example, compatibility judgments for the top items in the
\emph{memorable / unforgettable} scale were more similar than those for
\emph{good / excellent} or \emph{liked / loved.} Finally, there was
substantial consistency in ratings for items that were repeated across
experiments, suggesting that this paradigm elicited stable judgments
from participants.

@@

Figure \ref{fig:exp2_altsPlot} an example alternative set for the scalar
items \emph{liked} and \emph{loved} (combined). Alternative
distributions for the other scalar pairs (e.g.. \emph{good/excellent},
\emph{memorable/unforgettable}) were similarly long-tailed.

@@

Experiment 3 was conducted to measure pragmatic judgments
\(p_{L_1}(m \mid u)\) for five pairs of scalar items taken from Tiel
(2014). We include two conditions in Experiment 3. In the
two-alternative entailment condition, participants made judgments for
items included in the entailment scales. In the four-alternative
condition, participants made pragmatic judgments for the entialment
items and also the top two alternatives elicited for each scale in
Experiment 2. Running two pragmatic judgment conditions allowed us to
rule out the potential effects of having a larger set of alternatives
during the pragmatic judgment elicitation and also provided two sets of
human judgments to compare with model predictions.
\footnote{Note that alternatives in the four-alternative condition were chosen on the basis of Experiment 2, which was run chronologically after the two-alternative condition; both pragmatic listener experiments are grouped together for simplicity in reporting.}
